# Investigate Data

The data consists of 100,000 observations of space taken by the SDSS (Sloan Digital Sky Survey). Every observation is described by 17 feature columns and 1 class column which identifies it to be either a star, galaxy or quasar.

1. obj_ID = Object Identifier, the unique value that identifies the object in the image catalog used by the CAS
2. alpha = Right Ascension angle (at J2000 epoch)
3. delta = Declination angle (at J2000 epoch)
4. u = Ultraviolet filter in the photometric system
5. g = Green filter in the photometric system
6. r = Red filter in the photometric system
7. i = Near Infrared filter in the photometric system
8. z = Infrared filter in the photometric system
9. run_ID = Run Number used to identify the specific scan
10. rereun_ID = Rerun Number to specify how the image was processed
11. cam_col = Camera column to identify the scanline within the run
12. field_ID = Field number to identify each field
13. spec_obj_ID = Unique ID used for optical spectroscopic objects (this means that 2 different observations with the same spec_obj_ID must share the output class)
14. class = object class (galaxy, star or quasar object)
15. redshift = redshift value based on the increase in wavelength
16. plate = plate ID, identifies each plate in SDSS
17. MJD = Modified Julian Date, used to indicate when a given piece of SDSS data was taken
18. fiber_ID = fiber ID that identifies the fiber that pointed the light at the focal plane in each observation

```{r Load Libraries}
library(tidyverse)
library(MASS)
library(ggforce)
```


```{r Load Data}
# Set working directory
knitr::opts_knit$set(root.dir = here::here())
data <- read.csv(file.path(here::here(), 
                               "data", 
                               "raw", 
                               "star_classification.csv"))
```


```{r Data Structure}
str(data)
```

```{r}
data %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```


```{r Summarize Numeric}
data %>% 
  select_if(is.numeric) %>% 
  summary()
```

```{r Summarize target}
data %>% 
  group_by(class) %>% 
  summarize(count=n())
```

```{r Plot distributions}
numeric_variables <- c('u', 'g', 'r', 'i', 'z', 'redshift')

data %>% 
  dplyr::select(u, g, r, i, z, redshift) %>% 
  mutate(across(where(is.numeric), ~na_if(., -9999))) %>% 
  ggplot() +
  geom_density2d(aes(x = .panel_x, y = .panel_y)) +
  geom_autodensity() +
  geom_point(aes(x = .panel_x, y = .panel_y)) +
  facet_matrix(vars(u:redshift), layer.upper = 1, layer.diag = 2)
```


```{r}
data %>% 
  dplyr::select(u, g, r, i, z, redshift, class) %>% 
  mutate(across(where(is.numeric), ~na_if(., -9999))) %>% 
  ggplot() +
  geom_density2d(aes(x = .panel_x, y = .panel_y, color=class), show.legend = TRUE) +
  geom_autodensity() +
  geom_point(aes(x = .panel_x, y = .panel_y, color=class), show.legend = TRUE) +
  facet_matrix(vars(u:redshift), layer.upper = 1, layer.diag = 2)
```

